Spring haul-out behavior of seals in the Bering and Chukchi seas

### 2024-02-18 analysis/paper/Letter_Response_to_Reviewers.qmd

\hfill \break

Dr. Prange,

First, please accept my apologies for the significant delay in returning
revisions on this manuscript. A number of life and work priorities interfered
with a more timely response schedule and I sincerely appreciate the flexibility
*PeerJ* extends to authors. We are commited to a quick turnaround on any future
revisions or tasks associated with this manuscript and we look forward to a
final publication in *PeerJ*.

While many of the comments and suggestions from reviewers were minor, we feel
they deserved appropriate time, consideration, and exploration. We hope our
responses here reflect that consideration and we are grateful to the reviewers
for their thoughtful comments. Lastly, we also used this opportunity to improve
the quality of the underlying telemetry data set and expand the reproducibility
of the analysis to meet or exceed expectations for open science, data
availability, and code availability.

@@

# Response to Reviewers

Here, we respond directly to each of the suggested edits and comments. These
responses are organized by reviewer. All responses are prefaced with **\[author
response\]** and line numbers in the original submission are used as reference.
\hfill \break

\hrule

## **Reviewer 1 (A. Fischbach)**

> *The authors have put substantial efforts into setting the context for their
> analysis, citing relevant studies. I would like to see a more direct approach
> in the framing of their problem: the improvements of abundance estimates
> required by the Marine Mammal Protection Act and Endangered Species Act. I
> would like to see a specific example of how the modeled availability would
> have affect population abundance estimates by apply the availability model to
> aerial survey data and providing an adjusted abundance estimate.*

@@

::: author-response
**\[author response\]** We address this in the manuscript by introducing the
importance of the statutory requirements of the MMPA and ESA early in the second
paragraph of the introduction. We hope this frames the important link in the
mind of the reader between the work described here and the need for improved
abundance estimates.

Regarding the desire for specific examples of how the modeled availability would
have impacted population abundance estimates from aerial survey data, we don’t
have the ability to provide a full example because the aerial survey data are
not yet published and available. However, we note that Conn and Trukhanova
(2023, Marine Mammal Science 39:648-661) used some of our preliminary estimates
to assess potential aerial survey bias in using telemetered samples that are
biased towards younger age-sex classes. Their analysis is based on bias(N) =
bias(1/a) where a is availability (i.e., proportion hauled out).
:::

> *The findings are appear robust and intuitive. The authors need to improve the
> presentation of the findings to improve clarity and help the reader understand
> how they addressed the knowledge gap.*

@@

::: author-response
**\[author response\]** The reviewer provided numerous comments and suggestions
annotated throughout the PDF that address the concern of presentation and
clarity. In the next section we provide responses to those annotations
:::

### **Reviewer 1 (A. Fischbach) - PDF Annotated Comments & Suggested Edits**

> ***Title:** The reviewer suggests removing and aerial survey detectability
> from the title of the manuscript and suggests adding the seasonal time period
> our study focused on.*

@@

::: author-response
**\[author response\]** We have changed the title to: Spring haul-out behavior
of seals in the Bering and Chukchi seas
:::

> ***Abstract:** Can you make a statement in your abstract that use of modeled
> haulout behavior using covariates available from regional reananalysis data
> improved the confidence intervals on availability estimates and that this
> would improve aerial survey abundance estimation efforts?*

::: author-response
**\[author response\]** Unfortunately, we can't make this statement because 
inclusion of additional covariates (such as weather products) in haul-out models
only serves to increase standard errors in predictions.  The benefit of
including weather covariates is to more accurately explain variation in survey
counts; it is thus more relevant to reducing bias in aerial survey estimates
than it is to increasing precision.
:::

> *Please use direct language. You are not making findings on how seals rely on
> sea ice, rather you are improving methods to estimate their availability for
> improved aerial survey population estimation efforts. Here you are focused on
> modeling ice sea availability to aerial survey so that aerial survey counts
> may be appropriately interpreted. Please be clear that you are considering ice
> seal behaviors that may be telemtered as "dry out of water behaviors" that may
> be interpreted as behaviors that place the seal on top of the ice and
> available for observation by an aerial survey.*

@@

::: author-response
**\[author response\]** The opening sentence of the abstract is a statement
regarding the reliance of ice-associated seals on sea-ice for a range of
important life history behaviors. We think this statement is an important
introduction for readers who are not familiar with ice-associated seals. We do
agree with the reviewer's suggestion for more direct language and have made
edits to the beginning of the abstract with this in mind.
:::

> ***Abstract:** Did you use previously published data, or did you collect it? 
> Be specific using an active voice so that readers may understand your effort.*
>
> *For better flow, please indicate the behavioral data logging time scale here
> when you describe the data collection, not later when you speak to the
> analytical construct.*

@@

::: author-response
**\[author response\]** We have edited this to simply state *data from
satellite-linked bio-loggers* ... we have also altered this sentence to
specifically mention the hourly timescale for the bio-logger data
:::

> *The term "basking" implies a subset of activities that ice seals may engage
> in when hauled out on sea ice and not an activity that can be directly
> telemetered using your data logger.  Yet, what you are describing is "haul-out
> behavior".  Please use this term here.*

@@

::: author-response
**\[author response\]** edited to remove use of 'basking'
:::

> *Can you help your naive reader understand why you consider these covariates
> to be of interest and why others were not considered.*

@@

::: author-response
**\[author response\]** The main body of the paper includes this discussion.
Given the limited space and purpose of the abstract we feel it best to simply
list the covariates of interest
:::

> ***Line 41:** Please use specific terms and not metaphors so that
> second-language English readers may better understand your writing.*

@@

::: author-response
**\[author response\]** We have edited 'ripple effects' to 'tangible effects'
:::

> ***Line 41:** What about human communities that don't live in the region, but
> that send large LNG tankers and other vessel through the region?  The loss of
> sea ice has effects on those communities and activities too.*

::: author-response
**\[author response\]** Clearly there are a large number impacts of loss of
sea ice to human communities (and ecosystems) beyond the Arctic region to the
rest of the world. However, for this sentence, we wanted to keep the reader
focused on the Arctic region.
:::

> ***Line 44:** Please use direct language you mean biologically important
> functions?*

@@

To ensure the reader is clearly aware of which life-history behaviors/functions
we are referring to we provide specific examples within the sentence.
:::

> ***Line 50:** What does "readiness" mean in the context of a marine mammal?*

::: author-response
**\[author response\]** This is a good catch and we have edited the sentence to,
now, read:

"Knowledge about evolutionary constraints on the timing of reproductive and
molting behavior is generally lacking, so it is difficult to predict how or if
ice-associated seal species will adjust to future changes (e.g., by adjusting
pupping or molting schedules to earlier dates or different locales)"
:::

> ***Line 56:** Perhaps you mean to offer a topic sentence, stating "How
> ice-assoicated seals use sea ice as a haulout platform varies betweens
> species".*

@@

::: author-response
**\[author response\]** We have adopted this suggestion
:::

> ***Line 66-68:** There is so much to state about ice-associated seals in the
> Pacific Arctic that this reference to Antarctic work seems out of place here.*

@@

::: author-response
**\[author response\]** We have moved this reference to later in the paragraph
so the first references are about Arctic work. While the example is from the
Antarctic, it is an important, early example of analysis using bio-logger data
to estimate haul-out behavior in seals.
:::

> ***Line 77:** suggested edit*

::: author-response
**\[author response\]** We have adopted the suggested edit
:::

> ***Line 84:** Do you mean age-class? The term life-history implies knowledge
> of previous behavioral efforts or occupancy of regions of interest.   Here you
> are specifically interested in age-class.*

@@

::: author-response
**\[author response\]** We have edited to use the term 'demographic'
:::

> ***Line 92:** Please help your reader understand why are you doing your
> analysis and not offer a wide and jargon rich inventory of concepts that may
> be considered.*

> ***Line 93-94:** Please keep your introduction focused on establishing the
> knowledge gap that you are seeking to fill. Please do not lead your reader to
> the numerous other knowledge gaps that are not being addressed by your current
> study.*

@@

::: author-response
**\[author response\]** We have removed the sentence in question
:::

> ***Line 99:** Please use the term "hauled-out". Neither the biologger nor the
> aerial survey is able to distinguish among the various seal behaviors that
> result in them being hauled out and available for view by an aerial survey.*

@@

::: author-response
**\[author response\]** noted, and the terms "hauled out", "to haul out", etc is
used consistently throughout the manuscript
:::

> ***Line 114:** You have indicate that you would consider sea ice extent. Here
> you introduce the concept of icescape. Please introduce this concept earlier
> and indicate how you will incorporate this into your analysis.*

@@

> ***Line 117:** Please use an active voice for clarity.*

::: author-response
**\[author response\]** This sentence has been changed to:

"For this study we used haul-out behavior data and location estimates from
bio-loggers deployed on bearded, ribbon, and spotted seals in the Bering,
Chukchi, and western Beaufort seas by multiple organizations as part of
collaborative investigations from 2005 through 2020."
:::

> ***Line 120:** You summarize the years during which land-based studies applied
> bio-loggers.\
> Please do the same for the offshore deployments.*

@@

::: author-response
**\[author response\]** This sentence now reads:

"Ship-based capture events occurred during spring near the southern ice edge in
the Bering Sea between 2005 and 2018."
:::

> ***Line 124:** Please provide the geographic context of your tag deployments
> and data collection using a plot similar to that used by Jay et al. (2012:
> 10.3354/meps10057 figure 2). Please consider using a tabulation that enables
> readers to understand the geographic context of the deployments similar to
> that used by Jay et al. 2012, table 1.*

@@

::: author-response
**\[author response\]** We feel the spatial distribution of the actual
transmitted data -- compared to the deployment location -- is most informative
and relevant for the audience's understanding of how our data are distributed
spatially and how well data represents the range of each species. We provide
this visually in Figure 3 and adding an additional layer to that figure to show
deployments would be an additional complication. To ensure the audience has a
clear understanding of our study area, we have added an additional figure (now,
Figure 1) with a map of the Bering and Chukchi sea region that indicates
generally where deployments were initiated. In response to other reviewer
comments this figure also indicates the spatial range of aerial survey effort to
date.
:::

> ***Line 126:** Please use active voice for all actions that you took in your
> data collection, preparation and analysis.*

@@

::: author-response
**\[author response\]** Noted and addressed
:::

> ***Line 129-135:** Please move this to the discussion. This is a summary of
> your data findings, not a description of your methods.*

@@

::: author-response
**\[author response\]** We can see the reviewer's point. However, we do not feel
this topic is a summary of our data findings in this paper. Instead, our view is
that this information provides the context and rationale for our methodological
choice regarding bio-logger type and attachment methods. It is also important
regarding the analytic methods -- preference for flipper tag haul-out data over
those from a tag adhered to the dorsal/head. If we were to move these important
points to the discussion, we worry those points would be left disconnected.
:::

> ***Line 136:** This is a discussion of your methods. This is not your methods.
> Please state how the data was collected in your Methods section and discuss
> how data collection have been uncertain or biased in your discussion section.*

@@

> ***Line 141-143:** Please move to your discussion.*

::: author-response
**\[author response\]** We appreciate this suggestion from the reviewer and we
have moved large portions of this paragraph to the discussion. However, some
information regarding age/sex class determination is left here so the reader has
context for Table 1.
:::

> ***Line 144:** Please indicate how this was determined during data
> collection.*

@@

::: author-response
**\[author response\]** We have edited this sentence some to clearly communicate
that field determination of reproductive status (and, thus ideal delineation of
sub-adult and adult age class) was not possible.
:::

> ***Line 147-148:** Please allow your readers the opportunity to understand the
> geographic context of these deployments through use of symbols that maybe used
> to cross-reference the deployment locations in a map that indicates deployment
> locations along with filtered movement tracks.*

@@

::: author-response
**\[author response\]** As previously addressed, there are now two figures
available that provide geographic context of the deployments. Figure 3 shows the
geographic distribution of the data from the deployments. The paper is not
presenting or discussing movement ecology and, therefore, we do not present
filtered movement tracks. We keep the focus on haul-out behavior data.
:::

> ***Figure 1:** Please consider bar plot similar to that used by Jay et al.
> (2012: 10.3354/meps10057 figure 3), replacing the region with the seal
> species.*

@@

::: author-response
**\[author response\]** This figure has been updated with this suggestion in
mind
:::

> ***Line 151:** Please state clearly how you determined the end of a
> deployment. Here you indicate various elements of the haulout chronology,
> movement and dive chronology that may have been used, however not reader could
> replicate your methods base don this text.*

@@

::: author-response
**\[author response\]** We've added the following text to the paragraph:

"For example, a data record that ends with several consecutive days (\~10+ days)
of 100% dry observations and with locations indicating the tag was on land would
be truncated to the final stretch of 100% dry observations. The vast
majority of deployments end with the device detaching in the water and the
deployment end date is obvious. There is no perfect algorithm for identifying
deployment end dates and each deployment in question must be considered
separately. While not perfect, we are confident our reliance on expert judgement
and examination of multiple data streams provides the best option."
:::

> ***Line 152:** Deployment activity is known to disrupt normal pinniped
> behavior. Did you censure some period of data collection during and after the
> deployment to ensure that that disruptions from the deployment activity did
> not bias your analysis?*

@@

::: author-response
**\[author response\]** For these phocid species, we have no empirical data (or
even anecdotal observational data) on how captures may or may not disrupt
'normal' behavior. This is likely to be influenced by the capture method, how
long the animal was restrained, whether the animal was with a dependent pup,
whether the animal was molting, environmental variables such as wind and
temperature, etc etc. Given all these unknowns, we have not applied a censure to
the beginning of our deployment data. We can't 100% ensure this choice doesn't
result in some level of bias. However, given the large sample size and duration
of deployments we expect any bias imposed is negligible.
:::

> ***Line 154:** This is a result. Please indicate the method that you used to
> distinguish "plausible" from "implausible" telemetered seal behavior. In your
> Results section, you may indicate how method was able to identify occasions of
> implausible seal behavior and any patterns that arose relative to tag model or
> attachment.*

@@

::: author-response
**\[author response\]** This sentence has been moved to the results section.
Additionally, we've provided a brief explanation of implausible along with a
reference for more details. For these devices, the percent-dry timeline reported
100% dry for several weeks followed by an additional period of time that was
more typical of expected bearded seal behavior. During the extended dry time the
locations indicated movement of the tag that would not be simply drifting ice.
We also observed increasing transmission rates. The working hypothesis (i.e.
unconfirmed guess) is that the wet/dry sensor surface became obscured by some
corrosive element such that when in the water conductivity was prevented and,
thus, recorded as dry.
:::

> ***Line 156-164:** Please move to the Results section. Please move the
> discussion of how molt timing affects termination of tag deployments to the
> discussion section.*

@@

::: author-response
**\[author response\]** This figure has been moved to the results section along
with additional descriptive text. We have left the discussion regarding molt
timing and termination of the data record here to better communicate the point
to the reader and prevent an isolated mention in the discussion that has no
anchor.
:::

> ***Line 169-170:** Thus, the a successful transmission of data was required at
> no greater than 48 h intervals.*

@@

::: author-response
**\[author response\]** We're not sure how to interpret this comment. The main
point of the sentence is to explain the rationale/benefits of transmitting the
limited selection of percent dry values. We have expanded the text to provide
more details regarding the caching of messages and data transmission.
:::

> ***Line 174:** Do you mean "atmospheric" magnetic interference.\
> Presumably "extra-terrestrial atmospheric" events (such as events on Venus or
> Mars) played no role in the transmission of the data signal to the Argos
> receivers in orbit around Earth.*

@@

::: author-response
**\[author response\]** We have edited to the reviewer's suggested word
'encompasses'
:::

> ***Line 197:** Please use the active voice.*

::: author-response
**\[author response\]** Thank you for catching this. We have edited this
sentence to active voice
:::

> ***Line 198-199:** For most tags, your daily locations are from the Argos
> system, not from your data loggers. However, you included some FastLoc tag
> deployments. Please re-write for clarity of what these location are.*

@@

::: author-response
**\[author response\]** The previous paragraph describes how we determined daily
locations. We see the reviewers point re: the data not technically coming from
our data loggers. We have edited the sentence for additional clarity.
:::

> ***Line 199:** What is a "key" weather variable? Please write for clarity.*

::: author-response
**\[author response\]** We've edited this sentence and removed the word 'key'
:::

> ***Line 203:** This is a procedural detail. Other readers may replicate your
> method without needing to do this. Please omit this detail.*

> ***Line 207-208:** Readers can replicate your methods without the details on
> converting the NARR wind vectors into wind speed.*

::: author-response
**\[author response\]** We're not sure why the reviewer would recommend omitting
a 'procedural detail' as part of our methods. We feel it is important to clearly
communicate our procedure for processing the large NARR data set and that we
used the original grid size of 32km. Regarding the calculation of wind vectors, we
include this detail because many readers, like us initially, may not be well
versed in how wind vectors components are converted to wind speed.
:::

> ***Line 210:** Please use an active voice: "For all seal species we considered
> the following variables when modeling the hourly haulout behavior..."*

::: author-response
**\[author response\]** We have adopted the reviewer's suggestion
:::

> ***Line 224:** Using the term "major focus" implies that there may be numerous
> minor foci. Can you not state that this is the goal of the study? Ancillary
> natural history insights can be discussed incidental to this, but are not
> driving the research. The statutory requirements of the MMPA and ESA are
> driving your study. These Federal laws require robust methods to monitor
> population status. Your study seeks to build this method. Discussion of
> natural history belong in your Discussion section and not in the framing of
> your study. Please help the reader focus on why the Federal government is
> doing this study.*

> *Understanding pinniped haulout behavior is essential to designing and
> interpreting population estimation aerial surveys. Understanding how they use
> sea ice during a time when sea ice availability is rapidly changing is also
> important, but is ancillary to the paper's focus.*

@@

::: author-response
**\[author response\]** We have removed the term 'major focus' and, as the
reviewer suggests, simply state that this is the goal of the study. We have also
added mention of the statutory requirements to paragraph 2 of the Introduction
section and expanded the text to acknowledge the importance of understanding how
seals use sea ice during this period of change but, ultimately, it's ancillary
to the focus of this paper.

As for providing the reader focus on why the Federal government is doing this
study: this is a collaborative study across multiple organizations and the
analysis/purpose is not limited to the Federal government. We've incorporated
the reviewers suggestions to expand discussion of the statutory requirements of
the MMPA and ESA but don't feel we need to single out the motivations for the
Federal government.

We have also added additional general representation of recent aerial survey
effort to the new Figure 1. This should better communicate to the reader how the
spatial distribution of our haul-out behavior overlaps with the species range
and survey effort.
:::

### 2024-02-18 analysis/paper/London_HauloutBehavior.Rmd
final commit for revision

knitr::opts_chunk$set(dev = "ragg_png")

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  include = FALSE,
  fig.path = "../figures/",
  dpi = 300,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "100%",
  tab.topcaption = TRUE
)

@@

Again, we want to acknowledge Anthony Fisbach for the suggestion during the
peer review process. We think this is an excellent example of the peer review
process working to improve the quality of our manuscript and advance the
scientific process. We hope others will take our example and expand on it
within future analyses.

>>>>

knitr::opts_chunk$set(dev = c("ragg_png"))

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  include = FALSE,
  fig.path = "../figures/",
  dpi = 600,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "100%",
  tab.topcaption = TRUE
)

@@

Again, we want to acknowledge Anthony Fischbach for the suggestion during the
peer review process. We think this is an excellent example of the peer review
process working to improve the quality of our manuscript and advance the
scientific process. We hope others will take our example and expand on it
within future analyses.
